### 2020-10-12 presentation/presentation.tex
tex: ~fin

\ifcase 0  % choose 0=slides, 1=article, 2=refart
   \documentclass[aspectratio=169,ignorenonframetext,9pt]{beamer}
\or\documentclass[a4paper,11pt]{article}
   \usepackage{url,beamerarticle}
\or\documentclass[a4paper,11pt]{refart}
   \let\example\relax
   \usepackage{url,beamerarticle}
\fi

\ifcase 0  % choose a theme like these
    % \usetheme{boxes}
    \usetheme{Boadilla}
    % \usetheme{Goettingen}% I recommend
    % \usetheme{Singapore}
    % \usetheme{Pittsburgh}
    % \usetheme{Madrid}
    % \usetheme{Warsaw} % common choice, but often poor
\fi

\usepackage{algorithm,algorithmicx}
\usepackage{graphicx,pgfplots,parskip}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,url,array}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]
\newtheorem{algo}{Algorithm}[section]

\title{BSTERGM: Bayesian Separable-Temporal Exponential family Graph Models}
\author{Choi Seokjun}
\date{16 Oct. 2020}

\begin{document}

\begin{frame}
\maketitle
\end{frame}

% \begin{abstract}
% This abstract, being outside the frame environment, does not appear in the presentation.  Your outline will be the basis for a couple of sentences of talk for each of the following questions:
% \begin{itemize}
% \item What was done?
% \item Why do it?
% \item What were the results?
% \item What do the results mean in theory and/or practice?
% \item What is the reader's benefit?
% \item How can the readers use this information for themselves? 
% \end{itemize}
% \end{abstract}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{Model definitions and related terminologies}
\begin{frame}{Random graphs}
    \begin{defn}[Random graphs and related terminologies]
    Let $\mathcal{B}=\{0,1\}$. For given $n\in \mathbb{N}$,
    \begin{itemize}
        \item The set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes (without weights for each nodes and edges.)
        \item Let $\Omega$ be an event set. We say $Y: \Omega \to \mathcal{Y}$ is a random variable for a graph, or a random graph.
        \item For a random graph $Y \in \mathcal{Y}$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$, \\
            satisfying $Y_{ij}=1$ if the edge is connected. Otherwise, $Y_{ij}=0$.
        \item If edges of $Y$ have directions, then $Y$ is called a directed graph. Otherwise, $Y$ is called a undirected graph.
    \end{itemize}
    \end{defn}
    Let me notate a realization of random graph by $y$ and its edges by $y_{ij}$ for $i,j=1,2,...,n$.

    Here is a remark. These are obvious that, for given $n\in \mathbb{N}$,
    \begin{itemize}
        \item $|\mathcal{Y}|=2^{n(n-1)}$ if $\mathcal{Y}$ is the set of all directed graphs not permitting self-connecting edges.
        \item $|\mathcal{Y}|=2^{n(n-1)/2}$ if $\mathcal{Y}$ is the set of all graphs of undirected one.
    \end{itemize}
    ;thus, the size of $\mathcal{Y}$ grows exponentially when n increases.
\end{frame}

\begin{frame}{ERGM: Exponential family Random Graphs Models}
    \begin{defn}[ERGM: Exponential family Random Graphs Models]
        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes and $Y$ be a random graph (on $\mathcal{Y}$.)
        Set a distribution on $\mathcal{Y}$ to
        \[P(Y=y;\theta) = \frac{exp(\theta^{T}s(y))}{c(\theta)}\]
        for some $\theta\in\mathbb{R}^p$,
        where $s(y)\in\mathbb{R}^p$ is a vector which is part of y's sufficient network statistics,
        \\ and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
        \\ Models which have a such form called the ERGM.
    \end{defn}

    \begin{defn}[TERGM: Temporal Exponential family Random Graphs Models]
        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
        Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) to
        \[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta) = \frac{exp(\theta^{T}s(y_t, y_{t-1}))}{c(\theta, y_{t-1})}\]
        for $\theta\in\mathbb{R}^p$ and with the first-order Markov assumption $P(Y_2,Y_3,...,Y_T|Y_1)=P(Y_2|Y_1)P(Y_3|Y_2)...P(Y_T|Y_{T-1})$,
        \\where $s(y_t, y_{t-1})\in\mathbb{R}^n$ is a part of sufficient network statistics,
        and $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$ is a normalizing constant.
        Models which have a such form called the TERGM.
    \end{defn}
\end{frame}

\begin{frame}{STERGM: Separable-Temporal Exponential family Random Graphs Models}
    \begin{defn}[STERGM: Separable-Temporal Exponential family Random Graphs Models]
        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
        Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) by following way:

        Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$.
        Likewise, let $\mathcal{Y}^-|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or sparse edges comparing to $y_{t-1}$.
        Next, for $Y_t^+: \Omega \to\mathcal{Y}^+|_t$, $Y_t^-: \Omega \to\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
        \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})},
        P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
        for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
        and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1})), c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.

        Then, defining operations $+,-$ on $\mathcal{Y}$ following the boolean algebra edgewise-ly, set $y_t$ to
        \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]

        Additionally, assume that
        \begin{itemize}
            \item The first-order Markov assumption: $P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})$
            \item The separability: the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$;
                thus, \(P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+)P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\)
        \end{itemize}
        Models which have a such form called the STERGM.
    \end{defn}
\end{frame}

\begin{frame}{BSTERGM: Bayesian STERGM}
    To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
    Then, the posterior of $\theta^+,\theta^-$ becomes
    \[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})} \]
    where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$.

    Remarks:
    \begin{itemize}
        \item We cannot compute the normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically because we need to sum up too many terms.
        \item The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
            Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample.
    \end{itemize}
\end{frame}

\begin{frame}{Network Statistics for BSTERGM}
    As a usual, $s(y_{t}^.,y_{t-1})$ would be chosen by a difference of network statistics, $s'(y_t^.)-s'(y_{t-1})$.
    A common candidates of $s'$ are:
    \begin{itemize}
        \item the number of edges
        \item node degree distribution (each order)
        \item edgewise shared partner distribution (each order)
        \item dyadwise shared partner distribution (each order)
        \item k-star distribution
        \item triangle distribution
    \end{itemize}
    In fact, a bundle of all order of node degree distribution and all order of edgewise shared partner distribution form sufficient statistics of a graph.
    Also, we can set $s'$ by a function of these statistics.
\end{frame}

\section{Model fitting algorithm}
\begin{frame}{Fitting the BSTERGM}
    Suppose that a sequence of graph samples $y_1,...,y_T$ is observed
    \\ (then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined,) 
    \\ and we want to fit the observation using BSTERGM.

    How do we find the posterior distribution of $\theta^+,\theta^-$?

    Note that an ordinary MCMC algorithm does not work in our situation, because the constant part remains when calculating the MCMC ratio.
    To solve this doubly-intractable constant problem, we should use more special MCMC technique, the exchange MCMC algorithm.
\end{frame}

\begin{frame}{Fitting the BSTERGM}
    Here is the algorithm of our main chain for the exchange algorithm.
    \begin{algo}[the main chain]
    Let $y_1,...,y_T$ be given. For $m=1,...,M$,
    \begin{enumerate}
        \item Propose candidates $\theta_*^+,\theta_*^-$ from $\epsilon(.|\theta_{m-1}^+,\theta_{m-1}^-)$.
        \item Select a lag $(t-1,t)$ randomly on $2 \leq t \leq T$.
        \item Generate an exchange graph $y_{ex} \in\mathcal{Y}|_t$ (with $y_{ex}^+, y_{ex}^-$) at the $\theta_*^+,\theta_*^-$.
        \item Calculate the exchange MCMC ratio $\pi$ at the lag,
            \[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+,\theta_*^-)}
                {P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+,\theta_{m-1}^-)}
                \frac{P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}{P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)}\]
        \item With probability $min(\pi,1)$, accept the proposal and put $(\theta_m^+,\theta_m^-) = (\theta_*^+,\theta_*^-)$.\\
            Otherwise, reject the proposal and put $(\theta_m^+,\theta_m^-) = (\theta_{m-1}^+,\theta_{m-1}^-)$.
    \end{enumerate}
    \end{algo}    
    Observe that, the $\pi$ has no normalizing constant terms because they are canceled out by added exchange terms.
    \(log\pi = (\theta_*^+-\theta_{m-1}^+)(s(y_{t}^+,y_{t-1})-s(y_{ex}^+,y_{t-1}))
    +(\theta_*^- -\theta_{m-1}^-)(s(y_{t}^-,y_{t-1})-s(y_{ex}^-,y_{t-1}))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
    \( = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
    +(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
\end{frame}

\begin{frame}{Fitting the BSTERGM}
    We need one more thing. To generate the exchange graph $y_{ex}$ for time $t$ (of the second part of main chain), 
    we should have a generative algorithm at a given parameter points.
    However, we still do not know the normalizing constant. Thus, I use one more MCMC chain.
    \begin{algo}[the auxiliary chain]
    Let $\theta^+,\theta^-,y_{t-1}$ be given. For $k=1,...,K$,
    \begin{enumerate}
        \item Select one edge randomly from the $y_{k-1}$, say, $y_{k-1;ij}$.
        \item Propose a new graph $y_*$ with switching the $y_{ij}$ value from $y_{k-1}$: 
        \\ If $y_{k-1;ij}=1$, then set $y_{*;ij}=0$(dissolution case.) Oterwise, If $y_{k-1;ij}=0$, then set $y_{*;ij}=1$(formation case.)
        \\ (If sample graphs are undirected, switch $y_{ji}$ simultaneously.)
        \item Take $\theta$ as $\theta^+$ or $\theta^-$ according to the case. Calculate the MCMC ratio $\phi$,
        \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= \frac{exp(\theta^T s(y_*,y_{t-1}))}{exp(\theta^T s(y_{k-1},y_{t-1}))} = exp(\theta^T (s'(y*)-s'(y_{k-1})))\]
        \item With probability $min(\phi,1)$, accept the proposal and put $y_k=y_*$.\\
            Otherwise, reject the proposal and put $y_k=y_{k-1}$.
    \end{enumerate}
    \end{algo}
    After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
\end{frame}

\begin{frame}{MCMC Diagnosis}
    Since we run two kinds of MCMC chains, we should proceed two diagnostic task.

    The main chain produces the posterior samples of parameter, so the procedure is same as ordinary MCMC case.
    \begin{itemize}
        \item Cut burn-in period. Do thinning if it is needed.
        \item Depict traceplots of each parameter chain to check the convergence and the mixing.
        \item Depict autocorrelation plot. Calculate ESS if it is needed.
    \end{itemize}

    Next, checking all auxiliary chains is practically irritating one. 
    In general, it is suffice to check the auxiliary chain of the last iteration (of the main chain)
    with statistics included in the model.
    \begin{itemize}
        \item Calculate network statistics of all graphs produced by the last auxiliary chain.
        \item Depict traceplots of each statistics to check the convergence and the mixing.
    \end{itemize}
\end{frame}

\section{Model inference}
\begin{frame}{Inference and Prediction}
    Since we have posterior samples by running the main chain as many as we want, we can do basic inference procedure.
    \begin{itemize}
        \item A outlining shape of the posterior of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histogram.
        \item An approximated summary statistics: mean, mode, variance, ...
        \item An approximated quantile and probability interval
    \end{itemize}

    Moreover, we already have a generative algorithm at the specific parameter point,
    we can predict the form of network at $T+1,T+2,...$ using posterior sample following standard Bayesian method.
    For example, for predicting $T+1$,
    \begin{itemize}
        \item Run K iteration using auxiliary chain at $y_T$ with each posterior sample points.
        \item Take the each network as a predicted result.
    \end{itemize}
    If you need, calculate some network statistics for the results and get a summary statistics of them.
\end{frame}

\begin{frame}{Goodness of Fit}
    To evaluate the goodness of fit for the posterior $\theta^+,\theta^-$,
    we can use the auxiliary chain algorithm once again.
    \begin{algo}[GOF procedure]
        For t=2,...,T
        \\For s=1,...,S
        \begin{enumerate}
            \item sample $\theta_s^+,\theta_s^-$ from the estimate of posterior.
            \item simulate $y_s$ using the auxiliary chain under $y_{t-1}$.
            \item calculate $g(y_s)$, some higher degree statistics (eg. Node-degree dist \& Edgewise Shared Partner dist)
        \end{enumerate}
        Draw the box-plot of $g(y_s)$ and compare with $g(y_t)$.
    \end{algo}
\end{frame}

\begin{frame}{Supplements}
    You can find the C++ implementation (using Armadilo: see http://arma.sourceforge.net/)
    \\ of BSTERGM fitting, diagnostic, and GOF algorithms
    \\ at my Github page: https://github.com/letsjdosth/BayesianSTERGM.
\end{frame}

\end{document}